Trail Saving on Backtrack

### 2020-02-28 paper.tex
minor changes

\paragraph{Reason quality.}
The saved trail can be thought of as remembering the solver's recent
trajectory. Sometimes we want to follow the past trajectory, but
perhaps sometimes we do not. In particular, when adding literals from
$\trailsave{}$ to $\trail$ we can examine the quality of the saved
reasons to see if they are worth using. Once we encounter a literal
with a low quality saved reason we stop adding literals from
$\trailsave$ to the solver's trail. In particular, we can change
lines~\ref{ln:addToTrail} and \ref{ln:addToTrailNext} of \ust
to the following:\\
\begin{tabular}[t]{l}
23.5 \qquad \textbf{if} (lowQuality($\reasonsave{\litsave}$)) \textbf{break}\\
24.  \qquad $\trail.\mathrm{addToEnd(\litsave)}$\\
25.  \qquad $\reason{\litsave}\gets \reasonsave{\litsave}$
\end{tabular}

@@

\section{Experiments and Results}
We implemented our techniques in two different SAT solvers, MapleSAT and Cadical,
both of which have finished at or near the top of SAT competitions for the past several
years \cite{Heule2018ProceedingsOS,Heule2019ProceedingsOS}. We then ran each solver on the 800 total benchmark instances used in the main tracks of the 2018 and 2019 SAT competitions. The
experiments were executed on a cluster of 2.7 GHz Intel cores with 5000 seconds CPU time and 7 GB memory limits for each instance. We chose not to output or verify the proofs generated by any of the solvers. The Par-2 scores obtained and total instances solved by each solver are reported in Figures ~\ref{fig:cadicalNew}, ~\ref{fig:cadicalOld}, and ~\ref{fig:maple}.

@@

The version of cadical used in Figure~\ref{fig:cadicalOld} benefited from \cbt, which agrees with previously published results \cite{DBLP:conf/sat/MohleB19}. Trail saving alone did not significantly impact this solver, but adding all of the enhancements on top of trail saving resulted in solving the same number of instances as the solver with \cbt did.

MapleLCMDist (in Figure~\ref{fig:maple}) is another solver that benefited from \cbt, and trail saving alone solved more instances than the solver with \cbt did. Adding the first two enhancements on top of trail saving resulted in solving three more instances and yielding a better Par-2 score than the solver with \cbt did. Adding the last enhancement of ceasing trail saving on a "low quality" reason made the performance worse, whether clause size or lbd was used as the clause quality metric. This suggests that the enhancements to trail saving have different impacts on different solvers.

@@

\begin{figure}
\centering
    \begin{tabular}{|l|l|l|l|l|l|}
      \hline
      & Total & SAT & UNSAT & Avg. Par-2 \\ \hline
      cadical                  & 532          & 314 &  218  & 4025                             \\ \hline
      cadical-chrono          & 525          & 308 &  217  & 4086                             \\ \hline
      cadical-trail           & 529          & 310 &  \textbf{219}  & 4060                   \\ \hline
      cadical-trail-multipleBT & 530 & 313 & 217 & 3930 \\ \hline
      cadical-trail-multipleBT-lookahead & 531          & 313 &  218  & 4020                      \\ \hline
      cadical-trail-multipleBT-lookahead-reason & \textbf{538} & \textbf{319} & \textbf{219} & \textbf{3854} \\ \hline
    \end{tabular}
    \caption{Table of results for cadical, version pulled from github as of January 1, 2020.}
    \label{fig:cadicalNew}
\end{figure}

@@

\section{Conclusion}
We have shown that our method trail saving can speed up two
state-of-the-art \sat solvers, cadical and MapleSAT, about as well as
chronological backtracking can without breaking the conventional invariants
of CDCL. We also introduced three enhancements one can
implement when using a saved trail and demonstrated experimentally
that these enhancements can sometimes improve a solver's performance by a significant
amount. We have shown that trail saving and all enhancements we proposed are sound.
A line of future work might include introducing further enhancements,
such as using the saved trail to help make inprocessing techniques faster or
using the saved trail to learn a second learnt clause from a single conflict.
It is also possible to combine trail saving with chronological
backtracking if one is willing to break the traditional invariants of CDCL,
but we leave it to future work to determine whether or not this would be
useful and how to best approach it.

>>>>

\paragraph{Reason quality.}
The saved trail can be thought of as remembering the solver's recent
trajectory. Sometimes we want to follow the past trajectory, but
perhaps sometimes we do not. In particular, when adding literals from
$\trailsave{}$ to $\trail$ we can examine the quality of the saved
reasons to see if they are worth using. Once we encounter a literal
with a low quality saved reason we stop adding literals from
$\trailsave$ to the solver's trail. In particular, we can change
lines~\ref{ln:addToTrail}--\ref{ln:addToTrailNext} of \ust
to the following:\\[5pt]
\begin{tabular}[t]{l}
23.5 \qquad \textbf{if} (lowQuality($\reasonsave{\litsave}$)) \textbf{break}\\
24.  \qquad $\trail.\mathrm{addToEnd(\litsave)}$\\
25.  \qquad $\reason{\litsave}\gets \reasonsave{\litsave}$
\end{tabular}\\[5pt]

@@

\section{Experiments and Results}
We implemented our techniques in two different SAT solvers, MapleSAT
and Cadical, both of which have finished at or near the top of SAT
competitions for the past several years
\cite{Heule2018ProceedingsOS,Heule2019ProceedingsOS}. We then ran each
solver on the 800 total benchmark instances used in the main tracks of
the 2018 and 2019 SAT competitions. The experiments were executed on a
cluster of 2.7 GHz Intel cores with 5000 seconds CPU time and 7 GB
memory limits for each instance. We chose not to output or verify the
proofs generated by any of the solvers. The Par-2 scores obtained and
total instances solved by each solver are reported in
Figures~\ref{fig:cadicalNew}, \ref{fig:cadicalOld}, and
\ref{fig:maple}.

@@

\begin{figure}[t]
\centering
    \begin{tabular}{|l|l|l|l|l|l|}
      \hline
      & Total & SAT & UNSAT & Avg. Par-2 \\ \hline
      cadical                  & 532          & 314 &  218  & 4025                             \\ \hline
      cadical-chrono          & 525          & 308 &  217  & 4086                             \\ \hline
      cadical-trail           & 529          & 310 &  \textbf{219}  & 4060                   \\ \hline
      cadical-trail-multipleBT & 530 & 313 & 217 & 3930 \\ \hline
      cadical-trail-multipleBT-lookahead & 531          & 313 &  218  & 4020                      \\ \hline
      cadical-trail-multipleBT-lookahead-reason & \textbf{538} & \textbf{319} & \textbf{219} & \textbf{3854} \\ \hline
    \end{tabular}
    \caption{Table of results for cadical, version pulled from github as of January 1, 2020.}
    \label{fig:cadicalNew}
\end{figure}

The version of cadical used in Figure~\ref{fig:cadicalOld} did show
benefits from \cbt in agreement with previously published results
\cite{DBLP:conf/sat/MohleB19}. Trail saving alone did not
significantly impact this solver, but adding all of the enhancements
on top of trail saving resulted in solving the same number of
instances as the solver with \cbt did, albeit with a slight increase
in the Par-2 score.

MapleLCMDist (in Figure~\ref{fig:maple}) is another solver that
benefited from \cbt. In this solver trail saving alone solved more
instances than the solver with \cbt did. Adding the first two
enhancements on top of trail saving resulted in solving three more
instances and yielding a better Par-2 score than the solver with
\cbt. Adding the last enhancement of ceasing trail saving on a "low
quality" reason made the performance worse, whether clause size or lbd
was used as the clause quality metric. This suggests that the
enhancements to trail saving have different impacts on different
solvers.

@@

@@

\section{Conclusion}
We have shown that our trail saving technique can speed up two
state-of-the-art \sat solvers, cadical and MapleSAT, as or more
effectively than chronological backtracking can without breaking the
conventional invariants of CDCL. We also introduced three enhancements
one can implement when using a saved trail and demonstrated
experimentally that these enhancements can sometimes improve a
solver's performance by a significant amount. We have shown that trail
saving and all enhancements we proposed are sound.

There are many avenues that can be pursued in future work, such as
using the saved trail to help make inprocessing techniques faster or
using the saved trail to learn a multiple clauses from a single
conflict. It is also possible to combine trail saving with
chronological backtracking if one is willing to break the traditional
invariants of CDCL, but it would require further work to determine
whether or not this would be useful and how to best approach it.
